### 2021-02-08 codecheck-paper.tex
related work section
(I think this will need further tweaking)

\documentclass[12pt]{article}
\usepackage{f1000_styles}
\usepackage[numbers]{natbib}
\usepackage{hyperref}
\hypersetup{colorlinks=false, pdfborderstyle={/S/U/W 1}, pdfborder=0 0 1}
\usepackage{url}
\usepackage{setspace}

\usepackage{booktabs}
\usepackage{doi} % turn DOIs into links

\usepackage{academicons} % for ORCID icon, via https://tex.stackexchange.com/a/534476/212961
\definecolor{idcolor}{HTML}{A6CE39}
\newcommand{\orcid}[1]{\href{https://orcid.org/#1}{\color{idcolor}\aiOrcid}}

\newcommand{\email}[1]{\href{mailto:#1}{#1}}

\usepackage{fontspec}
\usepackage{fontawesome}
\newcommand{\github}[1]{\href{https://github.com/#1}{\faGithub}}
\newcommand{\twitter}[1]{\href{https://twitter.com/#1}{\faTwitter}}

\onehalfspacing

\begin{document}
%\title{CODECHECK: An open-science initiative for the independent execution of computations underlying research articles}
\title{CODECHECK: An Open Science initiative for the independent execution of computations underlying research articles during peer review improves reproducibility}
\author[1,$\ast$]{Daniel~N\"{u}st \orcid{0000-0002-0024-5046} \github{nuest} \twitter{nordholmen}}
\author[2,$\ast$]{Stephen~J~Eglen \orcid{0000-0001-8607-8025} \github{sje30} \twitter{StephenEglen}}
\affil[1]{Institute for Geoinformatics, University of M\"unster, Germany; \email{daniel.nuest@uni-muenster.de}}
\affil[2]{Department of Applied Mathematics and Theoretical Physics, University of Cambridge, UK; \email{sje30@cam.ac.uk}}
\affil[$\ast$]{Contributed equally}
\affil[$$]{Public project for this article is at \url{https://github.com/codecheckers/paper}.}
\maketitle
\begin{abstract}
  The traditional scientific paper falls short of effectively communicating computational research.
  Since this issue is difficult to address due to lacking incentives, guidance, benefits, recognitions, and reproducibility, we come at this problem from a different angle:
  We propose a system by which the computational workflows underlying research articles are checked.
  The CODECHECK system uses open intrastructure and tools and can be integrated into review and publication processes in multiple ways.
  We describe these integrations along multiple dimensions (importance, who, openness, when).
  In collaboration with academic publishers and conferences, we demonstrate CODECHECK with 25 reproductions of diverse scientific publications.
  These CODECHECKs show that asking for reproducible workflows during a collaborative review can effectively improve executability.
  While CODECHECK has clear limitations, it may represent a building block in Open Science and publishing ecosystems for improving the reproducibility, appreciation, and, potentially, the quality of non-textual research artefacts.
\end{abstract}

\clearpage

\section*{Introduction}\label{introduction}

Many areas of scientific research use computations to simulate or analyse their data. These computations are complex and are difficult to explain coherently in a paper \citep{marwick_how_2015}.
To complement the traditional route of sharing research by writing papers,
there is a growing demand to share the underlying artefacts, notably 
code and datasets, so that others can inspect, reproduce or expand that work
(see Figure~\ref{fig:inverse}).
Some of the earliest proponents of this initiative were Buckheit and Donoho \cite{buckheit_wavelab_1995}, who coined what has been called \emph{Claerbout's claim} (extending on \citet{claerbout_electronic_1992}):
\emph{``An article about computational science in a scientific publication 
is \text{not} the scholarship itself, it is merely \textbf{advertising} of
the scholarship. The actual scholarship is the complete software development
environment and the complete set of instructions which generated the 
figures.''}

If researchers start sharing more artefacts, how might these artefacts be examined to ensure that they do what they claim?
For example,
although  scientific journals now require a data sharing statement that
outlines what data the authors have (or will) share, journals  implement this differently.
At one end of the spectrum, journals have been created to accept ``data papers''
(e.g.,~\href{https://www.nature.com/sdata/}{\emph{Scientific~Data}}, 
\href{https://essd.copernicus.org/}{\emph{Earth System Science Data}}, 
\href{https://rmets.onlinelibrary.wiley.com/journal/20496060}{\emph{Geoscience Data Journal}},
\href{https://bdj.pensoft.net/}{\emph{Biodiversity Data Journal}},
\href{https://openpsychologydata.metajnl.com/}{\emph{Journal of Open Psychology Data}},
\href{https://odjar.org/}{\emph{Open Data Journal for Agricultural Research}},
\href{https://openhealthdata.metajnl.com}{\emph{Journal of Open Health Data}});
these journals have established rigorous procedures by which
the data is validated according to standards in each field.
At the other end of the spectrum, many journals still allow authors to state ``Data available upon reasonable request''.
Authors, while possibly well intentioned at the time of writing the article, often cannot provide data when requested as data disappears over time \cite{Vines2014-hf}.

Given that no clear standards yet exist for sharing data, what hope might there be for sharing computer programs?
Both data and software are required to validate a
computational analysis; data can be seen as inert whereas
code requires an environment to be run in.
This makes software harder to share.
Our experience is that researchers offer several reasons for why code
is not shared, e.g.,~``there is no documentation'', ``I cannot maintain it'', or ``I do not want to give away my code to competitors''.
Our view is that sharing code, wherever possible, is good for the community and the individual \cite{Barnes2010-iv,markowetz_five_2015}.
Having code and data openly available and archived provides a valuable
resource for others to learn from, even if the code is broken or lacks documentation.
However, with a little effort, we believe that if an independent person can execute an analysis, this is worth documenting and that this  reduces the barrier to evaluating non-text research materials.
Just as data journals' validations of data and all journals' peer review  provides a ``baseline reassurance'', i.e.,~that a paper has been checked by someone with an understanding of the topic \cite{fyfe_mission_2019}, the same baseline could be provided for the workflow underlying a paper.
With this in mind, we have developed a set of principles and an example workflow that provides a pragmatic way of checking that a paper's code works.

Here we offer a thorough description of a process and its variations to integrate a much-needed evaluation of computational reproducibility\footnote{Following the Claerbout/Donoho/Peng terminology, see \cite{barba_terminologies_2018}.} into peer review, and we demonstrate its feasibility by means of 25 reproductions across scientific disciplines.
We call this system CODECHECK.

\begin{figure}
% emojis: https://github.com/googlefonts/noto-emoji
% cross mark: u274c
% check mark: u2714
  \centering
  \includegraphics[width=0.7\textwidth]{figs/rr.pdf}
  \caption{The inverse problem in reproducible research. The left
  half of the diagram shows the diverse range of materials used
  within a laboratory. These materials are often then
  condensed for sharing with the outside world via the
  research paper, a static PDF document. Working backwards from the
  PDF to the underlying materials is impossible. This prohibits reuse
  and is not only non-transparent for a specific paper but is also 
  ineffective for science as a whole. By sharing the
  materials on the left, others outside the lab can reproduce
  or build on this work.}
  \label{fig:inverse}
\end{figure}

\section*{What is a CODECHECK?}\label{what-is-a-codecheck}

\subsection*{Workflow and people}\label{workflow-people}

CODECHECK is best demonstrated by way of our example workflow, and later
we expand on the underlying principles. The workflow involves three
groups of people:
(1) the AUTHOR of a paper providing the code to be checked,
(2) the PUBLISHER of a journal interested in publishing the author's paper, and
(3) the CODECHECKER, who checks that the authors's code works.
The six-step workflow we have refined is shown in Figure~\ref{fig:workflow}.

\begin{figure}
% original figure source: https://docs.google.com/drawings/d/1tGOZFNZle-oE1Ynw_tLZFmNcv3wgxD0HPs5bXzDOrng/edit
% emojis: https://github.com/googlefonts/noto-emoji and https://github.com/twitter/twemoji
% emoji codes:
% book stack: 1f4da
% package: 1f4e6
% file drawer: 1f5c4
% disk: 1f4be
% document: 1f4c4
% phd: 1f393
% detective: u1f575
% laptop: 1f4bb
% folder: 1f4c1 1f4c2
% receipt: 1f9fe
  \centering
      \includegraphics[width=\textwidth]{figs/codecheck_overview.pdf}
      \caption{The CODECHECK example process implementation.
        Codechecker act as a detectives: They investigate and record,
        but do not fix issues.  Numbers in bold refer to the six steps
        outlined in the text.}
  \label{fig:workflow}
\end{figure}

\textbf{Step 1:} The author submits their manuscript along with the code and data to the PUBLISHER.
The code and data need not be openly available at this point.
However, in many cases the code and data may be published on a code hosting platform,
such as GitHub or GitLab. Ideally, the AUTHOR is expecting the CODECHECK and
prepares for it, e.g.,~by asking a colleague to attempt a reproduction or by 
following the PUBLISHER's guidelines for sharing computational workflows.

\textbf{Step 2:} The publisher finds a CODECHECKER to check the
code. This is analogous to the publisher finding one or more
peer-reviewers to evaluate the paper, except we suggest that the
CODECHECKER and the author talk directly to each other.

\textbf{Step 3:} The CODECHECKER runs the code, based on instructions provided by
the author. They check if some or all of the results from the paper can be
reproduced. If there are
any problems running the code, the CODECHECKER asks the AUTHOR for help,
updates, or further documentation.
The burden to provide reproducible material lies with the author.
The CODECHECKER then tries to run the code again.
This process iterates until either the CODECHECKER is successful,
or the CODECHECKER concludes the code can not be executed within the context of a CODECHECK.
As part of this process, the CODECHECKER could work entirely locally, relying on their own computing resources, or in the cloud, e.g.,~using the open MyBinder infrastructure \cite{jupyter_binder_2018} or alternatives, some of which are more tailored to scientific publications while others offer commercial options for, e.g.,~publishers (cf. \cite{konkol_publishing_2020}).
A cloud-based infrastructure allows for the CODECHECKER and author to collaboratively improve the code and enforces a complete definition of the computing environment; but, unless secure infrastructure is provided, e.g.,~by the PUBLISHER, this requires the code and data to be published openly online.
Note that the task of the CODECHECKER is to check the ``mechanics'' of the workflow not to evaluate its correctness. 
In the context of mathematics, Stodden et~al. \cite{stodden_setting_2013} distinguish between \emph{verification} and \emph{validation};
following their definition, a CODECHECK ensures verification
\footnote{Note this is a different use from that referred to as the \emph{Verification Reports} article type in the journal \emph{Cortex} \cite{chambers_verification_2020}.},
of computational results, i.e.,~checking that the code correctly produces the output it claims to create, but not a validation, i.e.,~checking that the code implements the right algorithm to solve the specific research problem.
Nevertheless, simply attempting to reproduce an output may highlight a submisison's shortcomings in meeting a journal's requirements (cf.~\cite{christian_journal_2020}) and may effectively increase transparency, thereby improving practices (cf.~\cite{nosek_scientific_2012}) even if the check does not go into every detail.

\textbf{Step 4:} The CODECHECKER writes a certificate stating how the
code was run and includes a copy of outputs (e.g.,~data, figures, or
tables) that were independently generated.  The certificate may
include further recommendations on how to improve the material.  The
free text in the certificate can describe what was really checked,
because each workflow is unique.  Since no specific tool or platform
should be required, such that no authors are excluded, it would be
futile for the CODECHECKER to use automation or fixed checklists.

\textbf{Step 5:} The certificate and auxiliary files created
during the check, e.g.,~a specification of a computing environment, data 
subsets or helper scripts, and the original code and data get deposited in
an open archive unless restrictions regarding data size, license or 
sensitivity apply.
Currently, the CODECHECKERs deposit the material on Zenodo themselves, but the PUBLISHER may complete this step after integrating CODECHECK into its review process.
A badge or other visual aid may be added to the deposit and the paper and
link to the certificate.
Although a badge simplifies the CODECHECK into a binary value and risks introducing confusion regarding the extent of the check, especially across disciplines and methods, a badge can provide recognition value and acknowledgement of the completed CODECHECK.
The badge and the actual check are incentives for undertaking the effort needed to provide a readily computationally reproducible workflow.

\textbf{Step 6:} The PUBLISHER can, depending on the timing, provide the
certificate to peer-reviewers or handling editors or publish it
and link between  certificate, paper, and any repositories 
through metadata. Currently, the CODECHECKER creates these connections on
Zenodo\footnote{
See ``related identifiers'' and ``alternate identifiers'' in the right-hand
sidebar of certificate \texttt{2020-025} \cite{cert-2020-025}.}.
The PUBLISHER also credits the CODECHECKER for their work by depositing
the activity appropriately in scholarly profiles, such as ORCID\footnote{
See peer review contributions in ORCID records at
\url{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}.}.
The PUBLISHER also ensures proper publication metadata, e.g.,~links from the 
certificate repository to the published paper or the original code repository.

\newpage

\subsection*{Variations}\label{variations}

\subsubsection*{Dimensions of CODECHECK workflows}\label{dimensions-of-workflows}

The example workflow leaves room for variations to this process.
Here we consider several dimensions in a space of possible CODECHECK workflows(Figure~\ref{fig:dimensions}), to represent relevant aspects for stakeholders.
These aspects touch on timing, responsibilities, and 
transparency and will be discussed in the following sections.
The options on each dimension have their own pros and cons and can,
for the most part, be freely combined.

\begin{figure}
% original figure source: https://docs.google.com/drawings/d/1q2EdW3Ad-IpcD-CeoR1eIFMJW2fF6OVmDhCtCUmLO1k/edit
  \centering
      \includegraphics[width=0.8\textwidth]{figs/codecheck_dimensions.pdf}
  \caption{The dimensions of implementing a CODECHECK process.}
  \label{fig:dimensions}
\end{figure}

\subsubsection*{When to do a CODECHECK and with what importance?}
\label{when-to-do-a-codecheck}

The time at which a CODECHECK is done and its ascribed importance are closely connected, so we describe the dimensions \emph{When} and \emph{Importance} together.
The earlier a CODECHECK happens in the publishing process, the more it can affect editorial decisions:
Is a paper published, sent back for revisions, or rejected?
Even earlier checks, i.e.,~a CODECHECK of a preprint, may help to improve the workflow itself, even before a publisher is involved.
As such, codechecking papers could be part of a preprint server's
policy or initiated by interested authors.

Publishers could introduce a CODECHECK as a \textbf{strict prerequisite}.
As this can reduce the workload of reviewers, such a check should occur early in the review process.
Yet, the later in the review process the check happens, the easier is it to allow bidirectional communication between the author and codechecker, e.g., because the author might already be notified of the paper's acceptance and may more freely share materials online closer to the paper's publication date.
A \textbf{pre-review} CODECHECK means editors could send a submission
for peer review only if it passes the check, or include the certificate in the submission package provided to peer-reviewers.
Reviewers may then judge the relevance of the computations for the results of the work, or follow journal guidelines regarding how strictly to apply the check's outcomes in their evaluation.

A CODECHECK may also be conducted in \textbf{parallel} to the scientific peer review.
This puts less burden on the turnaround time for the CODECHECK, yet it only makes the outcomes available during the final consideration by the handling editor.
The check could also be assigned after suggestion by a reviewer, which would remove the need for submissions to undergo a pre-review screening.
However, soliciting such a ``specialist review'' is much less
desirable than having a regular CODECHECK, thus avoiding the situation
in which some submissions get special treatment. In both cases, the
editor's decision could be based both on CODECHECK and peer-review reports.

A \textbf{post-acceptance} CODECHECK would have the smallest impact on
editorial decisions and may simply provide \textbf{extra merit} on top
of the submission's acceptance.  This is the least impactful solution
in which all material is still evaluated and the results of the check
are properly acknowledged, because the check can be completed before
publication of the paper.  The GIScience checks (see
below) falls into this category: By displaying a badge on the volume
and article landing pages, the AGILE conference highlights articles
whose reproducibility was confirmed.  Similarly, in
collaborations with journals, some GIScience articles were
checked whilst authors worked on revisions.

A CODECHECK may also be conducted \textbf{post-publication}, though
this requires an update to the article and article metadata to
reference the check so that readers can find the CODECHECK.  In
general, publishers hesitate to make such revisions to published
articles.  We do not prefer this option as it has the least impact on
current publishing practices and downplays the importance
of reproducible workflows for ensuring good scientific practice.

%The timing of a CODECHECK needs connecting to typical scholarly communication process.
%While ideas exist for radically changing how researchers can share their work
% \footnote{For example Octopus, \url{https://science-octopus.org/about}, or Hypergraph, \url{https://www.libscie.org/hypergraph}.},
% % mimosa is another platform idea, but no real prototype AFAICT:
% % https://projects.invisionapp.com/share/E9Z3RKE7W3P#/screens/436978924
% % https://docs.google.com/presentation/d/1V_K8hghgnvGfEtW7TdwtowTveZyy4Qs8rMemNyHx_Dg/edit#slide=id.p
% an evolutionary approach has value, and we do not yet suggest to completely eliminating research papers as a way of sharing work.
Enhancing existing processes with CODECHECKs allows communities to
gradually transition towards more open practices.
When integrating a CODECHECK into existing review and publication processes, the \emph{turnaround time} is crucial.
Depending on when and who conducts the check, it might be done quickly
or it might delay publication.
We found that a CODECHECK generally takes  2--5 hours, with some outliers on the higher end.
This time includes writing and publishing the certificate but excludes
actual computation time, some of which took days.
These efforts are comparable to the time needed to peer review a submission,
which aligns with the efforts some volunteer codecheckers are willing to
make.
Currently, there is considerable amount of communicating about the
process, especially regarding who publishes which document when, so that proper
cross-referencing between paper and certificate is ensured via persistent
identifiers.  When integrated into a peer review platform, this
handling of documents should become much more streamlined.

\subsubsection*{Openness, or ``Who knows who?''}\label{who-knows-who}

Anonymity is broadly discussed, especially in the push towards
open peer review as part of the Open Science movement 
(cf. \cite{ross-hellauer_guidelines_2019}).
Without taking a strong stance on this topic, our motivation 
behind CODECHECK
for higher transparency and reproducibility does indeed favour a more open 
review process.
However, anonymity can protect individuals
\cite{tennant_limitations_2020}, e.g.,~junior scientists.  
The negative effects of a signed review may be reduced if a CODECHECK is not relevant for a journal's decision to accept or reject, but that is, of course, not desirable when the goal is higher transparency and reproducibility.
Instead, CODECHECK is a technical process that should generally find fixable problems; it is not aimed at giving an opinion or identifying a faulty approach.
If passing a CODECHECK becomes mandatory, full transparency may need
revisiting as the relations between authors and codecheckers would fall under the
same social and community challenges as open peer review
(cf.~\cite{everythinghertz123}).
% previous sentenc inspired by https://osf.io/9cftx/

The technical nature of the check and the challenge of providing sufficient documentation is why we see great benefits in bidirectional communication between author and codechecker.
Instead of trying to
fix problems or guess the next step, the codechecker can ask the author to 
rework the documentation or update code.
Instead of struggling to provide perfect instructions and as a result possibly not sharing any code or data, the author can make a best effort to document sufficiently.
Authors and readers can profit from a codecheckers' experience and approach, as during the check they may create useful and instructive files, e.g.,~a machine-readable computing environment specification.
While communication between author and codechecker may be anonymised via
the publisher, it most likely only helps to protect the identity of
the codechecker, because code is hard to anonymise.
Therefore, the most effective and desirable situation for the
stakeholders is to hold a open and collaborative CODECHECK.
The contributions by the codechecker may even be integrated into
the code of the workflow and be acknowledged as code commits. This way, 
proper credit can be given within the research software development community.

\subsubsection*{Who does the CODECHECK?}\label{who-does-the-codecheck}

Just as with peer-reviewers, a potential codechecker should have the
right skills and availability to do the work.  Ideally, the
codechecker has a matching code \emph{and} domain expertise to the
paper, although a well-documented workflow should be executable by any
computationally-competent person. Naturally, the more
prerequisite knowledge the codechecker has, the less time is needed to
understand the goals and the mechanics of an analysis.  From our
experiences, the priority should be given to matching technical
expertise first, as lacking knowledge in setting up a computing
environment with a particular language or tool is much more of a
problem than assessing the outcome, e.g.,~comparing created figures
with the original, without an in-depth understanding of the science.
The depth of the check will mostly be driven by the time required and
expertise of the checker, though in general, we expect a CODECHECK 
to consider reproducbility of the results above performance of the code.

Codecheckers could be drawn from a regular pool of volunteer
\textbf{scientific reviewers}, or from a special group of
\textbf{reproducibility reviewers} via specific roles such as
\textbf{reproducibility editors}, or \textbf{editorial staff} with a
publisher.  A \emph{single codechecker} is sufficient to verify the
workflow since it is mostly a factual process and not require much
debate or opinion.  Code usually harbours systematic and repeatable
mistakes and is thereby more reliable and auditable than processes
controlled by humans \cite{tibav:42484}, e.g.,~in a laboratory.  If
however publication of the paper depends on the CODECHECK, a second
opinion may be required.

We also see a great opportunity to involve \emph{early-career researchers} (ECRs) as codecheckers.
ECRs arguably have a high interest in learning
about the latest tools and technologies, as they build up their own
expertise and specialisation.
CODECHECK offes a way for ECRs to gain insights into new research and
highlight the importance of reproduction.  
\emph{ReScience X}, a journal devoted to reproduction and replication experiments \cite{roesch_new_2020}, shares an interest in this combination.
ECRs are also often familiar with new technologies, thus also making them likely to author CODECHECK-ready manuscripts
\footnote{Cf. the 77\% of 141 registered reports that were submitted to leading Neuroscience journals by ECRs \cite{chambers_registered_2019}.}.
As ECRs are introduced to peer review as codecheckers, they may transition into the role of scientific reviewer over time.
Overall, we see several opportunities and benefits to setting up a new process for CODECHECKING with a clear commitment to openness and transparency, independent of the current peer review process (see \emph{Openness} dimension).

The codechecker could be a member of \textbf{editorial staff}; this is
the most controlled but also resource-intensive option.  Such a
resource commitment would show that publishers are investing in
reproducibility, yet this commitment may be hard for small publishers.
These codecheckers could be fully integrated into internal workflows.
Credit for doing the codecheck is also achieved, as it is part of
their duties.  By contrast, it is useful for researchers to be
publicly credited for their reviewing activity.  A regular review
may be listed in public databases (e.g.,~ORCID
\footnote{\href{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}}
or commercial offerings such as Publons
\footnote{\href{https://publons.com/}{https://publons.com/}} or
ReviewerCredits\footnote{\href{https://www.reviewercredits.com/}{https://www.reviewercredits.com/}});
a codechecker could be similarly listed.
% The number of scientific reviews is of course only a coarse indicator of an individual's community contribution, yet it is unclear if a regular reviewer, who conductsa regular scientific review and in addition also conducts a CODECHECK, should be credited twice.
% In this case, the published CODECHECK certificate and possibly the published peer review can clearly describe the possibly large efforts put into both.

The codechecker
community\footnote{\href{https://github.com/codecheckers/codecheckers/}{https://github.com/codecheckers/codecheckers/}}
has over 20 volunteers who signed up in the last year.  Their
motivations, mentioned in the registration
information\footnote{\url{https://github.com/codecheckers/codecheckers/labels/registration}},
include: supporting reproducible research and Open Science, improve
coding skills, gaining experience in helping scientists with their
code, encouraging a sharing culture, and learning from other people's
mistakes; many are also motivated simply by curiosity.  We see
benefits to an open shared list of codecheckers across journals rather
than a private in-house group, as this may allow for better matches
regarding expertise and more even sharing of the workload.  Overall,
this community will aim to ensure that CODECHECK is a viable option
for independent no-cost Open Access journals.

\section*{Core principles}\label{core-principles}

The workflow and variations outlined describe our current
views on how code could be checked.  They are not immutable, but
we believe the following core principles underpin our CODECHECK
process:

\textbf{1. Codecheckers record but don't investigate or fix.} \\
The codechecker follows the author's instructions to run the code. If
instructions are unclear, or if code does not run, the codechecker
tells the author. We believe that the job of the codechecker is not to
fix all these problems but simply to report them and await a fix from
the author.  The level of documentation that is needed for third
parties to reproduce a workflow is hard to get right, and
too often this uncertainty leads researchers to give up and not
document it at all.  The conversation with a codechecker fixes this
problem.

\textbf{2. Communication between humans is key.} \\ 
Some code may
work without any interaction (e.g., CODECHECK certificate \texttt{2020-013}
\cite{cert-2020-013}) but more often there are hidden
dependencies or installations that need adjusting for a particular system.
Allowing the (human) codechecker to communicate
directly and openly with each other should make this process as
constructive as possible. We feel that routing this conversation
(possibly anonymously) through a publisher would introduce delays and
reduce the chances for community building.

\textbf{3. Credit is given to codecheckers.} \\
The value of performing a CODECHECK is comparable to that of a peer
review, and it may also require a similar amount of time. Therefore,
the codechecker's activity should be recorded, ideally in the
published paper.  The public record can be realised by publishing the
certificate in a citable form (i.e.,~with a DOI, as conducted in the
community CODECHECK process), by listing codecheckers on the journal's
website or, ideally, by publishing the checks alongside peer review
activities in public databases.

\textbf{4. Workflows must be auditable.} \\
The codechecker should have sufficient material to validate the
workflow outputs submitted by the
authors. Stark~\cite{stark_before_2018} calls this
``preproducibility'' and the ICERM report \cite{stodden_setting_2013}
defines the level ``Auditable Research'' similarly.  Every community
may establish their own good practices or adapt generic concepts and
practical tools, such as publishing all building blocks of science in
a research compendium
\footnote{\url{https://research-compendium.science/}}.  A completed
check means that code could be executed at least once using the
provided instructions, and, therefore, all code and data was given and
could be investigated more deeply or extended in the future.  Ideally,
this is a ``one~click'' step, but achieving this requires particular
skills and a sufficient level of documentation for third
parties. Furthermore, automation may lead to people gaming the system
or reliance on technology, which can often hide important details.
All such aspects can reduce the understandability of the material, so
we estimate our approach to codechecking, done without automation and
with open human communication, to be a simple way to ensure long-term
transparency and usefulness.  We acknowledge that others have argued
in favour of bitwise reproducibility because, in the long run, it can
be automated\footnote{Konrad Hinsen on Twitter:
  \url{https://twitter.com/khinsen/status/1242842759733665799}}, but
until then we need CODECHECK's approach.

\textbf{5. Open by default and transitional by disposition.} \\
By default, unless there are strong reasons to the contrary
(e.g.,~sensitive data on human subjects), all code and data, both from
author and codechecker, will be made freely available before or when
the certificate is published.  Openness is not required for the paper
itself, to accommodate journals in their transition to sustainable
Open Access models.  The code and data publication should follow
community good practices.  Ultimately we may find that CODECHECKS are
subsumed within peer review.
%% By disposition, a CODECHECK workflow should
%% be designed (a) recognising shortcomings in asserting reproducibility
%% of submissions and (b) with an understanding of the complex situation
%% that led to this. However, establishing a CODECHECK workflow should
%% mirror initiatives to improve education around reproducibility of
%% methods.  The long term goal is for a specific community to reach a
%% state where the level of verification ensured by codecheckers becomes
%% part of peer review and a regular merit of all accepted papers.

\section*{Implementation}\label{implementation}

\subsection*{Register}\label{register}

We have a curated list of 25 certificates available at
\url{https://codecheck.org.uk/register}.
Table~\ref{tab:register} lists the certificates
along with citations to the original papers.
These CODECHECKs fall into three themes:
(1) classic and current papers from computational neuroscience,
(2) COVID-19 modelling preprints, and
(3) GIScience.

The first theme was the initial set of papers used to explore the
concept of CODECHECK.  The idea was to take very well known historical
articles from a domains of interest to one of the co-authors (SJE).
Our first CODECHECK (certificate number \texttt{2020-001}) was
performed before publication on an article for the journal
\emph{GigaScience}, which visusalized the outputs from a family of
supervised classification algorithms.

The second theme was a response to the COVID-19 pandemic, choosing
modelling papers to predict outcomes. The checks were solicited
through community interaction or by our initiative rather than
requested from journals.  Some certificates were since acknowledged in
the accepted papers
\cite{Davies2020-vj,kucharski_effectiveness_2020}. In particular, we
codechecked the well-known Imperial college model of UK lockdown
procedures from March 2020(certificate \texttt{2020-10}).  Contrary to
opinion in the mainstream media, the model results were reproducible
(CITE NOORDEN).

The third theme represents co-author DN's service as a Reproducibility
Reviewer at the AGILE conference series, where the \emph{Reproducible
  AGILE} Initiative \cite{reproducible_agile} independently
established a process for reproducing workflows at the AGILE
conference series \cite{nust_improving_2020}.  While using slightly
different terms and infrastructure (``reproducibility reports'' are
published on the Open Science Framework (OSF) instead of certificates
on Zenodo) AGILE reproducibility reviews adhere to CODECHECK
principles.  A few checks were also completed as part of peer reviews
for GIScience journals.

\begin{table}
  \footnotesize

  \centering

  \begin{tabular}{llp{12cm}}
    \toprule
    \textbf{Certificate} & \textbf{Research area} & \textbf{Description} \\ \midrule
    \texttt{2020-001} \cite{cert-2020-001} & Machine learning & Code for benchmarking ML classification tool checked post acceptance of manuscript and before its publication in \textit{Gigascience} \cite{Piccolo2020-lo}. \\
    \texttt{2020-002}  \cite{cert-2020-002} & Neuroscience & Code written for this project checked by second project member as demonstration using paper from 1997 showing unsupervised learning from natural images \cite{Hancock1992-mp}. \\
    \texttt{2020-003}  \cite{cert-2020-003} & Neuroscience &  Code written for this project checked by second project member as demonstration using classic paper on models of associative memory \cite{Hopfield1982-mz}. \\
    \texttt{2020-004}  \cite{cert-2020-004} & Neuroscience & Code written for this project checked by second project member as demonstration using classic paper on cart-pole balancing problem \cite{Barto1983-rg}. \\
    \texttt{2020-005}  \cite{cert-2020-005} & Neuroscience & Check of independent reimplementation of spike-timing-dependent plasticity (STDP) model \cite{larisch_re_2019} conducted as demonstration for this paper. \\ % Python ANNarchy
    \texttt{2020-006}  \cite{cert-2020-006} & Neuroscience & Check of independent reimplementation of a generalized linear integrate-and-fire neural model \cite{detorakis_re_2017} conducted as demonstration for this paper. \\
    \texttt{2020-007}  \cite{cert-2020-007} & Neuroscience & Check of independent reimplementation of analysing spike patterns of neurons \cite{hathway_re_2018} conducted as demonstration for this paper. \\
    \texttt{2020-008}  \cite{cert-2020-008} & COVID-19 & Code for modelling of interventions on COVID-19 cases in the UK checked at preprint stage \cite{davies-preprint-2020} and later published \cite{Davies2020-vj}. \\
    \texttt{2020-009}  \cite{cert-2020-009} & COVID-19 & Code for analysis of effectiveness of measures to reduce transmission of SARS-CoV-2 checked as preprint \cite{kucharski-preprint-2020} and later published \cite{kucharski_effectiveness_2020}. \\ % R, check mentiond in acks, but not linked
    \texttt{2020-010}  \cite{cert-2020-010} & COVID-19 & Code for analysis of non-pharmaceutical interventions (Report 9) checked as a preprint \cite{ferguson_report_2020}. \\ % CovidSim
    \texttt{2020-011}  \cite{cert-2020-011} & COVID-19 & Code for modelling of COVID-19 spread across Europe was provided by authors and checked while paper was in press \cite{flaxman_estimating_2020}. \\
    \texttt{2020-012}  \cite{cert-2020-012} & COVID-19 & Code for modelling of COVID-19 spread across the USA was checked as preprint \cite{unwin_report_2020} and later published \cite{unwin_state-level_2020}. \\
    \texttt{2020-013}  \cite{cert-2020-013} & Neuroscience & Code for analsis of rest-activity patterns in people wihout con-mediated vision was checked as a preprint \cite{Spitschan2020.06.02.129502} after direct contact with the authors. \\ % MATLAB
    \texttt{2020-014}  \cite{cert-2020-014} & Neuroscience & Code for analysis of pertubation patterns of neural activity was checked after publication as part of publisher collaboration \cite{Sadeh2020}. \\ % MATLAB
    \texttt{2020-015}  \cite{cert-2020-015} & Neuroscience & Code for a neural network model for human focal seizures was checked after publication as part of publisher collaboration \cite{Liou2020}. \\ % MATLAB
    \texttt{2020-016}  \cite{cert-2020-016} & GIScience & Code for models demonstrating the Modifiable Aral Unit Problem (MAUP) in spatial data science \cite{Brunsdon2020} was checked during peer review. \\ % R
    \texttt{2020-017}  \cite{cert-2020-017} & GIScience & Code for spatial data handling, analysis, and visualisation using a variety of R packages \cite{Bivand2020} was checked after peer review before publication. \\ % R
    \texttt{2020-018}  \cite{cert-2020-018} & GIScience & AGILE conference reproducibility report using a demonstration data subset with cellular automaton for modeling dynamic phenomena \cite{Hojati2020}. \\ % R
    \texttt{2020-019}  \cite{cert-2020-019} & GIScience & AGILE conference reproducibility report with subsampled dataset for reachability analysis of suburban transportation using shared cars \cite{Illium2020}. \\
    \texttt{2020-020}  \cite{cert-2020-020} & GIScience & AGILE conference reproducibility report using a container for checking in-database windows operators for processing spatio-temporal data \cite{Werner2020}. \\ % Python + PostGIS
    \texttt{2020-021}  \cite{cert-2020-021} & GIScience & AGILE conference reproducibility report checking code for comparing supervised machine learning models for spatial nominal entity recognition \cite{Medad2020}. \\ % Python
    \texttt{2020-022}  \cite{cert-2020-022} & GIScience & AGILE conference reproducibility report checking code for visualising text analysis on intents and concepts from geo-analytic questions \cite{Xu2020}. \\ % Python
    \texttt{2020-023}  \cite{cert-2020-023} & GIScience & AGILE conference reproducibility report on analysis of spatial footprints of geo-tagged extreme weather events from social media \cite{Owuor2020}. \\
    \texttt{2020-024}  \cite{cert-2020-024} & Neuroscience & Code for multi-agent system for concept drift detection in electromyography \cite{vieira_driftage_2020} was checked during peer review. \\ % R
    \texttt{2020-025}  \cite{cert-2020-025} & GIScience & Adaptation and application of Local Indicators for Categorical Data (LICD) to archaeological data \cite{carrer_application_2021} was checked after peer review before publication. \\ % R
    \\ \bottomrule
  \end{tabular}
  \caption{Register of completed certificates as of December 2020. An interactive version
  is available at \url{http://codecheck.org.uk/register}.
  }
  \label{tab:register}
\end{table}

\subsection*{Annotated certificate and check metadata}\label{annotated-certificate}

At the end of each CODECHECK, the codechecker writes a certificate
stating which outputs from the original article, i.e.,~numbers,
figures or tables, could be reproduced.  This certificate is made
openly available so that all readers of the article can see which
elements were reproduced and what limitations or issues were found.
The certificate links to code and data used by the codechecker,
allowing others to build on the work.  The format of the certificates
evolved during the project, as we have learnt to automate different
aspects of the certification, such as inserting figures into
certificate documents and managing metadata of a check.  The metadata
is stored in a machine-readable structured file in YAML, the CODECHECK
configuration file \texttt{codecheck.yml}\footnote{ See technical
  specification of the CODECHECK configuration file at
  \url{https://codecheck.org.uk/spec/config/latest/}.}.  The
configuration file enables current and future automation of workflows
and meta-analyses.

Figure~\ref{fig:cert} shows pages 1--4 (of 10) of a typical
certificate \texttt{2020-012} \cite{cert-2020-012}.  The referenced
article \cite{unwin_report_2020} described models of COVID-19 across
the USA.  Figure~\ref{fig:cert}A shows the certificate number and its
DOI, which points to the certificate and any supplemental files on
Zenodo.  The CODECHECK logo is added for recognition and to denote
successful reproduction.  Figure~\ref{fig:cert}B provides the key
metadata as a table generated from \texttt{codecheck.yml}; it names
the paper that was checked (title, DOI), the authors, the codechecker,
when the check was performed, and in which repository related material
is accessible.  Figure~\ref{fig:cert}C shows a textual summary of how
the CODECHECK was performed and any interesting findings.
Figure~\ref{fig:cert}D (page 2 of the certificate) shows the outputs
that were generated based on the MANIFEST of output files in the
CODECHECK.  It shows the file name (Output), the description stating
to which figure/table each file should be compared in the original
paper (Comment), and the file size.  Page 3 of the certificate,
(Figure~\ref{fig:cert}E) gives detailed notes from the codechecker,
here documenting what steps were needed to run the code and that the
code took about 17~hours to complete. Finally, page 4 of the
certificate shows the first output generated by the CODECHECK
(Figure~\ref{fig:cert}F). In this case, the figure matched figure~4 of
\cite{unwin_report_2020}.  The remaining pages of the certificate show
other outputs and the computing environment in which the certificate
itself was created (not shown here).

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/annotate-cert-crop.pdf}
  \caption{Annotated certificate \texttt{2020-012} \cite{cert-2020-012} (first four pages only).}
  \label{fig:cert}
\end{figure}

\subsection*{Tools and resources}\label{tools}

We use freely available infrastructure, GitHub and Zenodo, to run our
system.  The \texttt{codecheckers} \textbf{GitHub}
organisation\footnote{ \url{https://github.com/codecheckers}} contains
projects for managing the project website, the codecheckers community
and its discussions, code repositories, and the main register of
CODECHECKs. Both the website\footnote{
  \url{https://codecheck.org.uk/}} and the register\footnote{
  \url{https://codecheck.org.uk/register}} are hosted as GitHub
pages. The register is a single table in CSV format that connects the
certificate identifier with the repository associated with a specific
conducted CODECHECK. Each of these repositories, which currently can
be hosted on GitHub or Open Science Framework (OSF), contains the
CODECHECK metadata file \texttt{codecheck.yml}. The register further
contains a column for the type of check, e.g.,~community, journal, or
conference, and the respective GitHub issue where communications and
assignments around a specific check are organised. No information is
duplicated between the register and the metadata files. The continuous
integration infrastructure of GitHub, GitHub Actions, is used to
automate the register for viewing (HTML) and for integration with
other services (CSV, JSON).  \textbf{Zenodo} is our preferred open repository for
storing certificates. It mints DOIs for deposits and ensures
long-term availability of all digital artefacts related to the
project. The CODECHECK community on Zenodo
\footnote{\url{https://zenodo.org/communities/codecheck/}} includes
certificates and the regularly archived register
\cite{codecheck_register_jan2021}.

A \textbf{custom R package}, \texttt{codecheck}\footnote{
  \url{https://github.com/codecheckers/codecheck}}, automates
repetitive tasks around authoring certificates and managing the
register.  It includes scripts to deposit certificates and related
files to Zenodo using the R package \texttt{zen4R} \cite{zen4r} and
for the register update process outlined above.  Codecheckers can
ignore this package, and use their own tools for creating
and depositing the certificate.  This flexibility  accommodates
different skill sets and unforseen technical advances or challenges.

These tools and resources demonstrate that a CODECHECK process can be
managed without infrastructure costs on freely available platforms.
The automation should ensure that more checks can be conducted in the
future, though having the process be managed by a paid CODECHECK
editor or further automation, e.g.,~with a bot, to reduce turnaround
time. Our main resource requirements are the \textbf{humans} needed for managing
the project and processes and the codecheckers.  All contributions
currently rely on (partly grant-based) public funding and
volunteering.

\section*{Related work}\label{related-work}

The journal \emph{ACM Transactions on Mathematical Software (TOMS)} 
recently established a ``Replicated Computational Results'' (RCR) review process
\cite{heroux_editorial_2015}, where ``replicable'' is the same as our
use of ``reproducible''.

Fifteen RCR Reports have been published so far
\footnote{\url{https://search.crossref.org/?q=\%22Replicated+Computations+Results+\%28RCR\%29+Report\%22&from_ui=yes&page=2};
  search executed on 2020-12-10.}
% 15: https://academic.microsoft.com/search?q=Replicated%20Computations%20Results%20%22(RCR)%20Report%22&f=&orderBy=0&skip=10&take=10
and the process is being extended extended to the ACM journal
\emph{Transactions on Modeling and Computer Simulation
  (TOMACS)}\footnote{\url{https://dl.acm.org/journal/tomacs/author-guidelines}}.
The TOMS RCR follows CODECHECK principles 1-4, although our work was
independently developed of theirs.  Here, we discuss variants of
CODECHECK, while TOMS RCR realises one particular workflow for
reproducibility reviews for one particular journal. The TOMS editorial
\cite{heroux_editorial_2015} shares similar concerns about selection
of reviewers, as we discussed above. Unlike existing CODECHECK
certificates, the RCR reports, which are similar to CODECHECK
certificates, undergo editorial review, but this could eventually
become part of the CODECHECK processes at journals.  The publication
of the RCR report is seen as a reward for the efforts of the
reproducing person, while the potential for this motive to be a
conflict of interest is acknowledged.  TOMS also recognises reviewer
activity in a partnership with Publons\footnote{
  \url{https://authors.acm.org/author-services/publons}}.
%, but it is
unclear if the author of an RCR report receives both the report's
%authorship and the reviewer recognition. 
CODECHECK however specifically aims to give codecheckers
recognition as reviewers.  In our view, the reviewer role removes the
possible conflict of interest while keeping the public
acknowledgement.  Specific to the field of mathematics, the RCR is
also expected to apply a review of the software itself if the system
it runs on cannot be evaluated by an independent party.  The TOMS RCR
creators concur with the importance of communication, expect
collaboration between author and RCR reviewers, share the
considerations around reviewer selection, and also put trust in
reviewer judgement over numerical bit-wise perfection.  A key
difference is that for TOMS RCR, authors opt-in with an \emph{RCR
  Review Request} and the RCR reports are published in the TOMS
journal next to the actual papers.
%% SJE: think this could be tightened a bit, and some of it I don't
%% fully understand.

Several journals provide special article types for reproductions of
published papers.  \emph{Information Systems} has an
invited reproducibility paper track\footnote{
  \url{https://www.elsevier.com/journals/information-systems/0306-4379/guide-for-authors}.}.
\emph{Nature Machine Intelligence} recently introduced a new type of
article, the reusability report \cite{noauthor_research_2020}.
Inspired by the detailed and nuanced submissions to a reproducibility
challenge, the reusability report focuses on the exploration of
robustness and generalizability of the original paper's claims
\cite{noauthor_research_2020}. This answers the specific community's
challenges around computational reproducibility and also values these
kinds of contributions as independent publications, which goes beyond
the goals of CODECHECK.  The journal \emph{Cortex} has a special
article type \emph{Verification Reports}, which are actually about
replication of results and are very well designed/reasoned
\cite{chambers_verification_2020}.  In a similar vein, the CODECHECK
certificates could also be published as a special article type within
journals.

Going beyond article types and a specific journal, the journal
\emph{ReScience C} publishes only
replications\footnote{\url{https://rescience.github.io/faq/}}, also
requiring open code and replication by a third party. ReScience also
relies on free infrastructure (GitHub and Zenodo).

For research with very high stakes, where reproduction would be too weak and
post-publication replication possibly too late because of policy impact,
Benjamin-Chung et~al. \cite{benjamin-chung_internal_2020} propose
\emph{internal replication}.
A workflow that has undergone internal
replication would very likely be of high quality and relatively easy to check.
Similarly, internal CODECHECKs may be used, with the same limitations such
as group think \cite{benjamin-chung_internal_2020},
to ensure reproducibility before submission.
Such internal checks are professionalised in local reproduction services,
such as CISER R-squared\footnote{
\url{https://ciser.cornell.edu/research/results-reproduction-r-squared-service/}} (established/permanent),
%Uni Münster's Reproducible Research Support Service\footnote{
%\url{https://go.wwu.de/r2s2}} (pilot),
Oxford's code review network (OxCRN)\footnote{
\url{https://github.com/OxfordCodeReviewNet/forum}} (community, broader scope), or
Yale's YARD\footnote{
\url{https://isps.yale.edu/research/data/approach}} (staff for replications).

Gavis and Donoho \cite{gavish_universal_2011} propose a new discipline
and infrastructure for reproducible computational research with its
own tools and API. Their specific packaging format, provenance record,
and cryptographic \emph{Verifiable Result Identifier} would indeed
provide excellent reproduciblity. However, the system is also complex
and since its creation in 2011 we are not aware of any publisher using
it; also, the system is not open source itself.  In comparison,
CODECHECK is less powerful but also much more flexible and less
dependent on specific tools or infrastructure. If data and code are
deposited properly, i.e.,~very unlikely to disappear, then the
certificate's DOI is practically close to the cryptographic
identifier.

Another platform for publishing results of reproductions is
\emph{SciGen.Report}\footnote{\url{https://scigen.report/}}.  It is a
community-run independent platform to foster communication on
reproducibility.  People can report on fully, partiially, or
un-successful reproductions of articles after publication.

No approach besides CODECHECK spans across journals and is
intentionally designed to be adopted across journals or events and to
build a community of codecheckers. CODECHECK shares its
interdisciplinary nature with other community initiatives concerned
with reproducibility awareness, education, and support, such as
ReproHack\footnote{\url{https://reprohack.github.io/reprohack-hq/}},
Code Copilot\footnote{\url{https://twitter.com/Code_Copilot}}, or
Papers with Code\footnote{\url{https://paperswithcode.com/about}}.
The latter recently announced a collaboration with the preprint server
\emph{arXiv} on providing data and code supplements for machine
learning
manuscripts\footnote{\url{https://medium.com/paperswithcode/papers-with-code-partners-with-arxiv-ecc362883167}}
and runs a reproduciblity
challenge\footnote{\url{https://paperswithcode.com/rc2020}}.
% Papers with Code now also does more disciplines than just ML.
Likewise, different disciplines and journals provide reproducibility
checklists, e.g.,~sciene and engineering \cite{rosenberg_next_2020} or
GIScience \cite{nust_agile_2019}, which naturally share some aspects
while addressing particularities as well as addressing researchers
from different fields. Regarding the education and guidance for
authors, we see CODECHECK's role as referencing and linking
educational efforts and helpful material, not as creating and
maintaining such content.

\section*{Limitations}\label{limitations}

\textbf{Isn't CODECHECK what peer review should be doing already?} \\
On the surface, yes.
However, the general observation is that the
increasing number of researchers, submissions, and reviews to conduct
under the pressure to publish puts the scholarly publication system
under significant strain.
Miyakawa described how when he as an editor simply requested raw data from authors, this led to almost all of those papers in which data was requested to be withdrawn or rejected \cite{miyakawa_no_2020}.
Such basic checks to show that
sufficient data supports the results requires much effort from editors,
yet they only rely on the availability of data files and the content of the
paper.
These availability checks can be enhanced by having more complex CODECHECKs request the code and then execute it, because code is not inert but needs to be run in a specific environment.
This might fall within idealistic expectations of peer review,
but is far from common in reality.
Establishing a CODECHECK
process is largely an acknowledgement that peer reviewing practices
have been unable to adapt to the challenges of digitisation of data-based science.
The concept of a CODECHECK,
just as the concepts of reproducible research and Open Science,
may be transitional by nature. If the activities described here as
being part of a CODECHECK become the norm for scientific reviews and
publications, then the initiative will have succeeded. However, the
rapidly changing technology used in research leads us to believe that
a separate role of codechecker and the expertise it brings will be
beneficial in the long term.

\newpage

\textbf{Should CODECHECK requirements not be more demanding?} \\
CODECHECK by design does not require authors to provide (and sustain) an
eternally functional workflow and neither suggests a specific software stack
or practical approach even if such suggestions, e.g.,~research compendia or ``repro-packs''\footnote{See \url{https://research-compendium.science/} respectively \cite{barba_praxis_2018}.}, are very valuable.
Creating something that anyone can reproduce has been called a 
fool's errand\footnote{\url{https://twitter.com/DougBlank/status/1135904909663068165?s=09}}
and we tend to agree.
However, the package of data, code, and documentation
collaboratively created by authors and codecheckers is a snapshot of a 
working analysis that greatly increases the likelihood of a successful 
reproduction and the possibility that a workflow can be extended by third parties in the future, as long as they have access
to suitable resources and matching skill set.
Concrete implementations of CODECHECK workflows, especially for specific
disciplines, may reify much more helpful guidelines for authors on how to
create reproducibility packages.
The author-friendly ``low bar'' should not stay low forever, but cultural
change takes time and the encouragement and guidance that CODECHECK,
as part of the widely accepted peer review concept, can provide may
eventually allow the bar to be raised as high as possible, e.g.,~with
executable research compendia \cite{nust_opening_2017}, ``Whole Tales'' \cite{brinckman_computing_2018},
or continuous analysis \cite{beaulieu-jones_reproducibility_2017-1}.
However, considering that missing artefacts and lack of documentation
have repeatedly been identified as key barriers to reproducibility
(e.g.,~\cite{stagge_assessing_2019,nust_improving_2020}),
we would not underestimate the power of a simple check.
For example, ModelDB curation policies ensure that only one figure is manually reproduced \cite{mcdougal_reproducibility_2016}, but that has not limited the usefulness nor success of the platform.

It should be noted that a codechecker does not fulfil the same role as a
\emph{statistical reviewer}, as it is applied by some journals in the 
biomedical domain (cf.~\cite{petrovecki_role_2009,greenwood_how_2015}).
The statistical reviewer actually evaluates the appropriateness of
statistical methods \cite{greenwood_how_2015} and can support topical
reviewers if, e.g.,~complex methods or sophisticated variants of statistical
tests are applied \cite{petrovecki_role_2009}.
The codechecker may go equally deep into the review, if they have
the required expertise and the required time, 
but they are by no means expected to be able to provide
such a level of scrutiny. We can imagine a tiered workflow where a 
codechecker could, just as a conventional reviewer could, recommend a detailed
statistical review to the editor if they come upon certain issues while
examining the work from a different perspective.

An important distinction is that a codechecker does not conduct a 
\emph{code review}. Code reviews are without a doubt extremely valuable in 
terms of improving reproducibility and reusability, and their proponents 
even believe they can improve the science \cite{petre_code_2014}.
Code reviews, however, have quite different structural challenges and 
require even more resources. That said, a well-reviewed codebase is likely
to be easier to codecheck, and the awareness of high-quality code raised
through CODECHECK may lead to more support for code reviewing.
Initiatives and journals that conduct software reviews independent of 
a specific publication or workflow include
ROpenSci\footnote{\url{https://ropensci.org/}},
PyOpenSci\footnote{\url{https://www.pyopensci.org/}},
and JOSS\footnote{\url{https://joss.theoj.org/}}.
Furthermore, the codechecker's task list is intentionally not
overloaded with related concerns, albeit potentially important ones,
such as ensuring proper citation of data and software or depositing
data and software in suitable repositories. Nevertheless,
we expect codecheckers to often recognise these issues and highlight them.

\textbf{How are failures during checks handled?}
We have no established process for the case
in which a reproduction completey fails.
This may be a sign that our pre-selection of the examples was biased, since so far all our CODECHECKs succeeded.
In the case that a journal adopts CODECHECK for all submissions, the open 
question remains as what to do if a check fails.
A negative comment in a CODECHECK certificate or a failed check does
not necessarily mean the paper or the science is bad
(cf. discussion on negative comments in \cite{everythinghertz123}).
At the start, we doubt that publicly reporting failures (i.e.,~the code 
would not run) will increase overall reproducibility, and we recommend not
sharing that outcome publicly but sharing the latest draft of the 
certificate with the authors.
One must be careful how policies on disclosing failures may negatively impact the acceptance by authors, but also how efforts by volunteers may go unnoticed if no report is published.
The most appropriate way forwards seems to be to give private feedback for failures and public acknowledgement and encouragement for successes.
% eLife's new model (see https://osf.io/9cftx/) will publish reports evantually, e.g.,~after preprint has been accepted elsewhere
Rosenthal et~al. \cite{Rosenthal2016b} discuss such incentives for different actors around the
implementation of reproducibility. We see CODECHECK as one way for
organisations to invest in reproducibility by creating incentives
until reproducible computations become the norm.

\textbf{Who will pay for the compute time?}
For those papers that take significant compute
time (days, not minutes), it is unclear who will pay for it.
The openness of CODECHECK certificates could help to establish community
peer pressure, or recognition, as submitting authors can be evaluated on their engagement
as codecheckers. This may lead to a balance of the shared load to reproduce
computations.
However, one must carfully consider the sustainability of rerunning computations and the
environmental impact large calculations, such as training machine learning models, have.
The challenges of computational resources can also quickly be reduced when
authors provide a small sample or a synthetic dataset together with the
expected outputs for this small example, which can equally well demonstrate the 
usefulness and validity of their workflow.

\textbf{What about my proprietary software and sensitive data?}
Authors can currently provide code that requires
proprietary software. Given the prevalence of proprietary software in
some disciplines, e.g MATLAB, this seems like a pragmatic choice.
However, this prohibits publications from benefiting from open infrastructure for reproducibility
(cf.~\cite{konkol_publishing_2020,perkel_make_2019})
and requires the codechecker to have access to that particular software.
Non-open software also considerably hampers reuse, especially by researchers
from the global south. Therefore, allowing proprietary software is a compromise
that each implementation of CODECHECK should critically reconsider.

Solutions for proprietary and sensitive data exists.
Authors can provide synthetic data (cf. \cite{shannon_opening_2018}), some
data can effectively be redacted \cite{oloughlin_data_2015}, and publishers or
independent entities can provide infrastructure for sharing data and workflows
confidentially \cite{perignon_certify_2019} or with access to derived results
but not raw data \cite{shannon_opening_2018},
i.e.,~data enclaves \cite{foster_research_2018},
or domains of reproducibility \cite{harris_more_2017})

\textbf{Can't someone cheat?} Yes. We don't check that the code is correct or
sound science,
just simply that it runs. This ``mechanical'' test is indeed a very low bar.
Yet, by having the code and (raw) data openly deposited, interested parties can
then examine the code, and we hope that knowing the code will be open will
be an incentive for authors to ensure good quality.
It will also give opportunities
to future researchers, potentially with new methods, to look for errors.
This prospective evaluation is a much more effective way to capture cheaters than engaging in an arms race on tools, methods, or resources trying to catch malicious intent now with closed datasets and code; this approach is comparable to storing blood samples of sport champions today to possibly detect doping in the future with more advanced methods (cf. \cite{everythinghertz97}).
Another comparison that helped us define the scope of a CODECHECK is was that we think of the codechecker as forensic photographer, making sure all the details
are captured so that an investigator may, now or in the future, scrutinize
them.

\textbf{Who's got time for more peer review?} Agree.
The time one spends performing CODECHECKs will likely reduce the time an individual has to work as a reviewer.
However, a CODECHECK is different from peer review.
First, the \textbf{technical nature} of a CODECHECK can provide a more 
clearly defined scope and expectations and thereby time budget compared to conventional 
peer review. 
For example, authors are told what to provide and the codechecker can be told when to stop.
The option for the codechecker to ask for clarification or more 
material can greatly reduce the digging required and thereby
increase \textbf{efficiency}.
Second, the specific skill set of a codechecker allows for \textbf{different groups} to take part in the review process.
ECRs might be attracted to learn more about recent methods, peer review, and reproducibility practices.
Research software experts (Research Software Engineers, RSE\footnote{\url{https://en.wikipedia.org/wiki/Research_software_engineering}}) who might
not regularly be involved in writing or reviewing papers might be interested in increasing their connection with scholarly practices.
An extra codechecker may simplify the matchmaking an editor has to do
when identifying suitable reviewers for a submission, as technical and
topical expertise can be provided by different people who do not even have to be traditional peer reviewers.
In fact, the codechecker could even be a software professional employed by a publisher.
Third, recall that we expect CODECHECKs not to be anonymous and always
publicly deposited, which is common practice for reviews of few journals only.
With code and workflows, there also is a good chance that a codechecker's
feedback can directly impact and improve the author's 
work, which may make for a rewarding experience or even provide tangible contributions.
The public certificates and contributions provide \textbf{peer recognition}.
Fourth, we also found that the focus on the mechanics of a
workflow and the ability to interact with the authors can make
reproductions very \textbf{educational}. It also is a different role and, as such, could
be a welcome option for researchers to give back their time to the community.

While such benefits are also part of idealistic peer review, they are mostly hidden behind paraphrased anonymous acknowledgement.

\textbf{Do workflows need to be codechecked multiple times?}
If you check a preprint before peer
review or a submission to a journal with non-public review, it might
need checking again as it is modified during peer review.
This is somewhat unavoidable, and has happened to us (e.g.,~certificate \texttt{2020-012}
\cite{cert-2020-012} has been revised). This can also be desired tough, if
the interaction between author, reviewer, and codechecker led to
improvements. Since checking the manuscript the second time is likely
to be much less work than the first time, this is largely mitigated by
open and regular communication. The higher the reproducibility and
automation, the more aspects of a CODECHECK that may eventually be
supported by automation, though we believe the human codecheckers
should remain the decision makers. Automatically detecting whether results
are ``the same'' is a question intrinsically linked with each unique
process and quite hard to realise for graphical outputs, such as plots or
maps.
That is why we do not require results to be exactly the same for a
CODECHECK to pass but simply that the code runs and generates the output
files that the author claims. Stochastic simulations mean that often
we cannot get exactly the same results, and that even the same versions of libraries
can generate outputs that differ by operating system
\cite{Gronenschild2012-pp}.
While reproducibility practices can mitigate some of these problems, e.g., by using a seed, the flexibility of the human judgement is still needed.

\textbf{Shouldn't the next step be more revolutionary?}
CODECHECK's approach is to acknowledge the shortcomings around 
computational reproducibility and to iteratively improve the current
system of scholarly communication based on journals and conferences.
It remains to be proven whether this approach is welcomed broadly and whether 
involving the traditional publishing stakeholders helps to further the cause.
We have discussed stronger more drastic rules at length, such as only 
considering fully free and open source software, diamond Open Access journals or community-led no-fee journals,
and journals with collaborative/open peer review, but we eventually decided
against them on the level of the principles.
For the CODECHECK community process\footnote{\url{https://codecheck.org.uk/guide/community-process}}
and the volunteer codechecker community, these requirements are reconsidered.

We also have deliberated requiring the latest technologial solutions which can
support reproducibility (cf.~\cite{konkol_publishing_2020}), but we have 
decided to focus on the human interface and the judgement of experienced
researchers and developers as a more sustainable and powerful approach.
The structurelessness of the required data and code gives all types of research the freedom to adopt CODECHECK.
CODECHECK could be complemented by automated scoring (e.g.,~\cite{menke_rigor_2020}), yet automation and metrics bear new risks.
Note that the focus of the CODECHECK principles on code execution allows journals
and publishers to innovate on the financial model and peer review practices
at their own pace.

\section*{Conclusions and future work}\label{future-work-and-conclusions}

CODECHECK works!
We have created a considerable number of certificates to demonstrate it.
The creation of the certificates and the interactions with authors and
editors not only improved the principles and the workflow but also
confirmed the approach taken. This result corroborates findings from
similar evaluations of reproducible computational research in journals and
at conferences.
CODECHECKs increase transparency of the checked papers and can contribute
to building trust in scientific results.
% "science earns trust through transparency" > https://twitter.com/briannosek/status/1290240244101521408
The set of 
shared principles and common name, through recognition value,
will allow researchers
across publication venues to judge the level of scrutiny that results have
faced. CODECHECK requires direct acknowledgement of the contribution to the
scientific community in the role of codechecker, not indirectly via citations 
of reproductions or informal crediation and closed inner circles.
Forther, codechecking can be a worthwhile educational activity, especially for
ECRs.

Yet, CODECHECK harbours the same limitations as peer review
in general and is closely connected to larger disruptions and challenges
in scholarly communication
\cite{eglen_recent_2018,tennant_ten_2019,fyfe_mission_2019}, 
including the tensions between commercial publishing and reviewers' often
free labour, and a global pandemic that has jumbled up scientific publishing
and exposed a broader general audience to science, even preliminary 
results \cite{munafo_what_2020}.
It is clear that establishing CODECHECK-like systems can be 
impeded by and must be seen as interconnected with much larger issues in 
science, such as broken metrics or malpractice triggered by publication
pressure \cite{piwowar_altmetrics:_2013,nosek_promoting_2015}.
We certainly do not want the binary attribute of ``code works''
to become a factor in bibliometric approaches for performance assesments.
% Mittermaier, Peer Review and Bibliometrics: http://hdl.handle.net/2128/22745

While developed for the current ``paper''-centric publication process and
being compatible with different styles of peer review, the CODECHECK 
principles would also work well with novel publication paradigms, e.g.,
peer-reviewed computational notebooks \cite{earthcube_new_2020},
iterative and granular communication of scientific outputs, 
articles with live-code \cite{perkel_pioneering_2019-1}
\footnote{For example, eLife's ERA, see 
\url{https://elifesciences.org/labs/dc5acbde/welcome-to-a-new-era-of-reproducible-publishing}.},
decentralized infrastructure and public reviewer reputation systems
\cite{tenorio-fornes_towards_2019},
and completely new visions for scholarly communication and peer 
review\footnote{Cf., for example, \emph{A modern vision for peer review} by 
Amy J. Ko: \url{https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07}}.
An explicit segmentation of research steps could even make the focus 
of a CODECHECK easier by only checking the ``analysis'' sub-publication.
The discovery of CODECHECKs could be increased by depositing the reports
into public databases of reproductions, such as \emph{SciGen.Report}.
Public researcher profiles, such as ORCID, may consider different
types of reviewer activity to capture how independent code execution contributes to science.
Notably, the discussed limitations are largely self-imposed for easier
acceptance and evolutionary integration, as to not break the current 
system and increase demands gradually without leaving practitioners behind.

If adopted today, a CODECHECK system, even if temporary in a sustainable 
transition towards more open publication and review practices, can contribute
to increased trust in the results of research. Introducing CODECHECK
should be informed by lessons learned from (introducing) open peer review
(OPR) \cite{ross-hellauer_guidelines_2019}.
All our conversations with publishers and editors at some point
touched on the procedural and technical implications of openness, be it 
that OPR is neither applied by nor familiar to stakeholders, or that process
management software would not support such open communication.

More reproducible practices initiated by CODECHECKs could lead
communities or whole disciplines to reach a state where authors provide 
sufficient material and reviewers have acquired sufficient skills that
the reviewers will generally conduct a CODECHECK-level of checking;
only in especially sophisticated cases will a specialised codechecker
be needed.
The main challenge for us remains getting journals to embrace the
idea behind CODECHECK and to realise processes that conform to the principles,
whether or not they use CODECHECK by name.
We would be keen to use the flexibility of the principles and cooperate
with journals to learn more about the advantages and yet unclear
specific challenges -- Do CODECHECKs really work better with OPR?
To facilitate the adoption, the CODECHECK badge is, intentionally, not branded beyond the checkmark
and green colour and simply states ``code works''.

Each rollout of CODECHECK should be accompanied by studies to ensure 
codechecking does not fall into the same gaps as peer review did
\cite{tennant_limitations_2020}
and to ensure positive change within the review system.
This \emph{cultural change}, however, is needed for the valuation of the
efforts that go into proper evaluation of papers.
Multiple adopting journals will also help to answer relevant questions on
the establishing process: What are crucial decisions or pain points? Can
authors retract code/data once a CODECHECK has started?
What variants of CODECHECKs will be most common? How will open CODECHECKs
influence or codevelop with the scope and anonymity of conventional review over time.

With the implementation of CODECHECK workflows, the question of training
codecheckers will become relevant. We expect both a mentoring scheme within
the CODECHECK community (experienced codecheckers provide on-the-job training
or serve as fallback advisors) as well as collaboration with reproducible
research initiatives such as
ReproHack\footnote{\url{https://reprohack.github.io/reprohack-hq/}},
ReproducibiliTea\footnote{\url{https://reproducibilitea.org/}}
\cite{fitzgibbon_brewing_2020},
and Repro4Everyone\footnote{\url{https://repro4everyone.org/}}
\cite{auer_reproducibility_2020},
to be able to grow the pool of codecheckers as needed.
The initial reaction of the scientific community and the spontaneous signing
up volunteer codecheckers make us optimistic that the time is right for 
putting scholarly peer review on the path towards facilitating the sharing and execution
of computational workflows.

\newpage

\subsection*{Competing interests}

SJE was on the editorial board at the journal \emph{Scientific Data} until January 2021.
DN is reproducibility chair at the Association of Geographic Information Laboratories in Europe's (AGILE) annual conference.

\subsection*{Grant information}

This work was financially supported by the UK Software
Sustainability Institute and a Mozilla Science mini grant.
DN is supported by grant
\href{https://gepris.dfg.de/gepris/projekt/415851837}{PE~1632/17-1}
from the German Research Foundation (DFG).

\subsection*{Acknowledgements}\label{acknowledgements}
%Collaborators and potential collaborators:
% someone from SpringerNature

We are grateful to the following individuals for discussions regarding
the work presented here: Andy Collings, Melissa Harrison, Giuliano
Maciocci, Naomi Penfold, Emmy Tsang (eLife), Rebecca Kirk (PLOS Computational Biology), Scott
Edmunds (GigaScience), and Andrew Hufton (Scientific Data). Iain Davies and
Yuhao (Sebastian) Wang developed code and example certificates.
We thank Antonio P\'{a}ez (Journal of Geographical Systems) for enabling
CODECHECKs, Carlos Granell and Frank Ostermann for contributing
certificates as reproducibility reviewers at the AGILE conference, and 
all authors of auditable workflows for their participation.
We thank Celeste R. Brennecka from the Scientific Editing Service, University of M\"unster, for her editorial review.

\subsection*{Author contributions}

DN and SJE contributed equally to all aspects of this project.

{\small\bibliographystyle{unsrtnat}
\bibliography{bibliography}}

\end{document}

% LocalWords:  CODECHECK CODECHECKER Zenodo ORCID ECRs codecheck codechecker

% Local Variables:
% TeX-engine: xetex
% End: