### 2021-05-27 reviews/response-round1.md
rename

<https://doi.org/10.5256/f1000research.54932.r82470>

We thank the reviewers for their comments on our manuscript.  The
comments are below.  Indented are our responses.

## Reviewer 1

In this article, authors propose to implement a procedure to check for
the code accompanying a submission to a journal. To do so, they
describe a pipeline made of 6 steps that ultimately lead to the
delivery of a code check certificate meaning that someone external to
the author's lab has managed to re-run the code. At this point, no
checking that the results are correct is necessary. The authors
already issued several codecheck certificates in different
disciplines. I find the idea really nice and certainly necessary but
I've a few questions (even though some of them are already addressed
in the "limitations" section). Given the structure of the paper, I'll
just list my questions here:

1. How does CODECHECK compare to ACM Artifact reviews badges?
(https://www.acm.org/publications/policies/artifact-review-and-badging-current)

    These badges, introduced in August 2020, show whether code is
    available (different levels) and reproduces same results.  In
    principle CODECHECKER could award these badges (artifacts
    evaluated, functional).  We have made a note to this effect in the
    manuscript (Related work, end of paragraph 1.)

2. What would be the incentive for someone to code check the code? Being
aware of the increasing difficulty in finding reviewers, I don't think
it would be easy to recruit people to perform a task that can rapidly
become very technical and time consuming.

    This was addressed in our section "Who's got time for more peer
    review?" We would however note that we have a pool of about 20
    volunteers currently willing to do codechecks.

3. How do you handle the case when specific hardware is necessary
(e.g. NVidia GPU)? Is it documented somewhere such that code-checkers
might first verify if they have the necessary hardware to run the
code?

    This was handled in the limitation "What about my proprietary
    software and sensitive data." but we now mention hardware too in
    the first paragraph of that section.

4. How do you establish a check has failed? For example, what happens if
a code-checker gets a segfault (for some unknown reason) and the
author is unable to help. Is it deemed failed?

    We hope that codechecker and author can resolve problems, but in
    the end there may be problems that cannot be solved.  Open
    infrastructure could help as both author and codechecker can work
    together in the same environment to minimise these failures.
    Ultimately however, there may be failures, which are noted in the
    section "How are failures during checks handled?".

5. Who will pay for the computing resources needed to run heavy
simulations and/or to acquire necessary software such as e.g. Matlab?
When a simulation consumes a lot of resources, it might wise to give
the checker access to computing resources. This could be paid for by
the journal.

    In the section "Who will pay for compute time?" we mention this
	problem, and that toy examples might alleviate the need to re-run
	resource-intensive computations.  We agree that one model might be
	that a journal provide some resource for this service.  Likewise,
	in the following paragraph, we describe that our pragmatic
	approach for now is to find codecheckers that have access to
	particular software, e.g. MATLAB.

6. I did not see in the report example a description of the environment
necessary to run the software. How did you solve the "dependency
hell"? Since the code might break at some point in the future because
of incompatibility in some libraries or environments, it would be
necessary to have a mechanism describing the running environment such
that it can be re-run later.

    The short answer is "we didn't.".  In the paragraph "Should
    CODECHECK requirements be more demanding?" we note our low bar of
    simply getting a codecheck to run once.  We do, however, encourage
    CODECHECKERS to describe the environment in free text form in
    their report.  Moving towards machine-readable descriptions would
    be a natural extension.

7. What do you recommend if the reviews are both excellents but the code
check failed? Does this mean the paper is blocked until code check
passes or rejected or else?

   This is up to the editor of the journal -- see the "Importance"
   dimension of Figure 3.  At one end, it could indeed be a "strict
   requirement" to get a codecheck certificate for the paper to be
   accepted.  On the other hand, it could be entirely optional.

8. The code check proposal is close to some extents to the Journal of
Open Science Software where each reviewer is assigned a list of things
to check during the review. Do authors consider this pipeline when
establishing their own pipeline?

   We have not considered this pipeline, nor do we have an explicit
   idea.  We now note this reviewer list  at the end of the third paragraph
   of "Related Work".

9. To what extent the codecheck certificate can be updated automatically
via some kind of "manual continuous-integration"? I mean that when
reading a paper online, would it be possible to click a button to test
if the code still runs considering the latest versions of libraries?
(for example, the certificate has been issued for Python 2 but I want
to know if this is usable with Python 3).

    To follow on from point 6, this would make a natural extension, but for
    now we are still considering one point in time, and keeping the
    requirements as close to the authors as we can.

10. When you look at journals advertising open data policies, it is
unfortunately not rare to find articles in these same journals without
the actual data. Do you have some suggestion for educating editors to
actually enforce the code check a journal adopt it?

    We share this concern, and unfortunately have no simple
    suggestions for helping editors.  At this early stage, we think
    the approach should be one of encouraging uptake, rather than
    mandating it.  We also hope that having specific in-house
    experience, e.g. editorial staff to examine for code and data
    availability, can note this.  But at the end of the day, this
    again is dependent on the journal's workflow.

Some suggestions:

11. The badge that is delivered would need some time information since the
check is valid at one point in time (with a given software stack) and
does not guarantee future runs.

    Great idea. we could add the certificate number to the URL, or add
    the certificate number.  We will try to implement this when
    revising our workflows.

12. For specialized journals, you could consider to offer a common generic
environment where a code could be first tested. It this fails, then
you would need only to slightly modify the environment to add missing
dependencies. For example, in neuroscience, a Neuro Debian would
probably suit the needs of a large number of models.

    Yes. We will certainly bear this in mind in future work.

13. As editor-in-chief of ReScience C, I would like to inform authors
that the journal now accepts "reproduction report". The idea it to try
to re-run the code accompanying a published article and to report if
it succeeded or failed. Our own procedure to check for reproduction is
not standardized and we'll certainly benefit from the code check
initiative.

    Thank you for noting this.  We now mention the reproduction report
    in the manuscript where we describe Rescience C.

Overall, it's nice to have a clean description of a pipeline to check
for code even though some questions need to be addressed. Also, I'm
not too confident that journals will adopt it immediately and I'm
afraid such initiative will take time to be generalized. But we have
to start somewhere.

    We share your realistic assessment that (a) journals may be slow
    to adopt but that (b) we should start somewhere.

## Reviewer 2

1. Is there any community feedback on the transparency and
reusability of research published with CODECHECK certificates?

This is perhaps too big of an ask this early in the
initiative as research reuse and citations are independent factors of
the publication and peer review of this specific paper - but I’d still
be interested in any insights the authors have to offer on this topic.

    This is an excellent question.  It is perhaps too early for us to
	assess this given most certificates are under a year old, but we
	are not aware of any reuse yet of the codecheck-deposited
	material.  However, we certainly think it interesting to try and
	monitor this over a longer (3-5 year) timescale if possible.   As
	well as looking for citations of certificates, we could also check
	for forks of our repositories, and download statistics from
	Zenodo.